### 2024-09-24 paper/paper.qmd
Your commit message here

---
title: "Data analysis of dangerous dog attack"
subtitle: "Analyzing the circumstance of dog attack and compare data to find out the most dangerous breed"
author: 
  - Yuanchen Miao
  - Another author
thanks: "Code and data are available at: https://github.com/NevaeH-9/dangerous-dog-attack.git."
date: 24 september 2024
date-format: long
abstract: "Nowadays, almost every family choose to have a pet, mostly dogs. Dogs have their innate nature to protect themselves or their owner from the danger they meet, sometimes they bark at it, sometimes they make an attack on it. Dogs require training from human to prevent them from attacking or barking at people, but some breed of dogs have stronger aggressive desire, in some situations, these breed of dogs probably attack people after training. In this report, data of Registry of Dogs Subject to and Issued a Dangerous Dog Order from Open Data Toronto is used, this data is refreshed monthly to today and all died dogs are removed from the data. There appears some specific breed of dogs are more likely to attack people and causing worse attck circumstance."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
```

# Introduction {#sec-introduction}

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

\newpage

# References